---
title: "Descriptive Statistics"
output: 
  pdf_document:
    latex_engine: xelatex
    keep_tex: yes
  header-includes:
  - "\\usepackage{subcaption}"
mainfont: "Arial"
fontsize: 12pt
date: "2025-11-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)

library(tidyverse)
library(openalexR)
library(tidyverse)
library(here)
library(bibliometrix)
library(bibliometrixData)
library(ggwordcloud)
library(kableExtra)
library(readxl)
```

# 1. Data Overview

## 1.1 Description of the Data Gathering Process

The data considered in the project is from OpenAlex. we use a pair-based keyword search to identify articles which use or construct game-theoretical models. The keywords we are taken from two lists: 1. A list of JEL codes defining the subject of study, i.e. game theory. 2. A list of terms describing game setup which ensures that we only cover papers employing or constructing game-theortic models. For each pair of keywords, we identify articles based on the appearance of keyword pairs in the title or abstract of the paper (\color{red} maybe we also want to consider full texts? This would increase the number of articles drastically.\color{black}) Currently, we focus only on articles and reviews written in English (\color{red}Should we include books?\color{black}). We drop articles from the corpus without any reference information. As the number of articles is rather low prior to 1980, we choose this year as the starting point for our analysis. 


## 1.2 Search Terms

```{r terms, results='asis'}

jel_terms <- read_excel(here("JEL_code_terms.xlsx"))

game_terms <- read_excel(here("game_description_terms.xlsx"))

t1 <- jel_terms 
t2 <- game_terms 

kable(list(t1, t2),"latex", booktabs=T, valign="t", linesep="", caption = "List of Query Terms") %>%
  kable_styling(latex_options = "HOLD_position", font_size = 10)

```


```{r data load, cache=TRUE}

data_raw <- readRDS(here("data_final_1980.RDS"))

data_concepts <- data_raw |> 
    select(oa_id, concepts) |> 
    unnest(cols = c(concepts)) 

data <- readRDS(here("descriptives.RDS"))

fields_of_interest <- c("Computer Science", "Business, Management and Accounting", "Decision Sciences", "Social Sciences", "Engineering", "Economics, Econometrics and Finance", "Mathematics", "Psychology", "Physics and Astronomy")

```

\newpage


## 1.3 Descriptive Summary of the Data

```{r descriptives}
kable(data$description_table, booktabs=T, "latex", linesep="", caption = "Descriptive Summary") |> 
  kable_styling(position = "center", latex_options = "HOLD_position")
```

\newpage

## 1.4 Top Articles

```{r articles}
top_articles <- data$top_articles 
top_articles[13,1] <- NA
top_articles <- top_articles[1:15,]
top_articles <- top_articles |> 
  select(authors,title, publication_year, cited_by_count, field_1)
colnames(top_articles) <- c("Authors","Title", "Year","Citations", "Field")
kable(top_articles, booktabs=T, "latex", linesep="", caption = "Top Articles in Terms of Citations") |> 
  kable_styling(latex_options = c("center","HOLD_position"), font_size = 8) |> 
  column_spec(1, width = "5cm", border_right = ) |> 
  column_spec(2, width = "5cm") |> 
  column_spec(5, width = "3cm") 
```

## 1.5 Number of Articles per Query

```{r articles_per_query, out.width="100%", fig.height=16, fig.width=12}
data_raw |> 
    group_by(query) |> 
    summarise(N = n()) |> 
    ggplot(aes(x=fct_reorder(query,N), y=N)) +
    geom_col() +
    coord_flip() +
    labs(x = "N", y = "Query")
    
```





# 2. Publications over Time

```{r publications over time, fig.cap="Publication and Citations over Time"}
N_over_time_tbl <- data$N_over_time_tbl

ggplot(N_over_time_tbl, aes(publication_year, No_publications)) +
    geom_line() +
    labs(y="Publicatiions", x="Year") +
    theme_bw() +
    theme(panel.grid = element_blank())
  
ggplot(N_over_time_tbl, aes(publication_year, No_citations)) +
    geom_line()+
    labs(y="Citations", x="Year") +
    theme_bw() +
    theme(panel.grid = element_blank())
```

# 3. Domain Overview

## 3.1 Top Domains

```{r domains, fig.cap="Number of Publication or Citations per Domain or Field"}

articles_per_domain <- data$articles_per_domain


ggplot(articles_per_domain,aes(No_publications, domain_1)) +
    geom_bar(stat = "identity", color="black", fill = "grey") +
    labs(y = "Domain", x="Publications") +
    theme_bw() +
    theme(panel.grid = element_blank()) +
 scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))  # 0% below, 5% above
  )  
ggplot(articles_per_domain,aes(No_citations, domain_1)) +
    geom_bar(stat = "identity", color="black", fill = "grey") +
    labs(y = "Domain", x="Citations") +
    theme_bw() +
    theme(panel.grid = element_blank())  +
 scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))  # 0% below, 5% above
  )
  
```

## 3.2 Domains over Time

```{r domains over time, fig.cap="Number of Publications per Domain or Field over Time"}
N_publications_over_time_domain_tbl<- data$N_publications_over_time_domain_tbl
N_citations_over_time_domain_tbl <- data$N_citations_over_time_domain_tbl

N_publications_over_time_domain_tbl |> 
    pivot_longer(!publication_year, names_to = "Domain",values_to = "No_publications") |> 
    ggplot(aes(publication_year,No_publications, color=Domain)) +
    geom_line() +
    labs(x = "Year", y="Publications") +
    theme_bw() +
    theme(panel.grid = element_blank(), legend.position = "bottom") +
    scale_x_continuous(expand = c(0,0), limits = c(1980,2025))

N_citations_over_time_domain_tbl |> 
    pivot_longer(!publication_year, names_to = "Domain",values_to = "No_citations") |> 
    ggplot(aes(publication_year,No_citations, color=Domain)) +
    geom_line() +
    labs(x = "Year", y="Citations") +
    theme_bw() +
    theme(panel.grid = element_blank(), legend.position = "bottom") +
    scale_x_continuous(expand = c(0,0), limits = c(1980,2025))
  

```

# 4. Field Overview

## 4.1 List of Fields

```{r list of fields}

data$articles_per_field |> 
  select(field_1) |> 
  rename(Field = field_1) |> 
  kable("latex", booktabs = T, linesep = "", caption = "List of Fields by OpenAlex.") |> 
  kable_styling(latex_options = "hold_position")

```

\newpage

## 4.2 Top Fields

```{r, fields, fig.height=10}
  
articles_per_field <- data$articles_per_field 

f1 <- ggplot(articles_per_field,aes(No_publications, fct_reorder(field_1, No_publications))) +
    geom_bar(stat = "identity", color="black", fill = "grey") +
    labs(y = "Field", x="Publications") +
    theme_bw() +
    theme(panel.grid = element_blank()) +
 scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))  # 0% below, 5% above
  )  
f2 <- ggplot(articles_per_field,aes(No_citations, fct_reorder(field_1, No_citations))) +
    geom_bar(stat = "identity", color="black", fill = "grey") +
    labs(y = "Field", x="Citations") +
    theme_bw() +
    theme(panel.grid = element_blank()) +
 scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))  # 0% below, 5% above
  )
gridExtra::grid.arrange(f1, f2, nrow=2)
```


## 4.3 Fields over Time


```{r fields over time}
N_publications_over_time_field_tbl<-data$N_publications_over_time_field_tbl |> 
  select(any_of(fields_of_interest))
N_citations_over_time_field_tbl <- data$N_citations_over_time_field_tbl |> 
  select(any_of(fields_of_interest))

N_publications_over_time_field_tbl |> 
    pivot_longer(!publication_year, names_to = "field_1",values_to = "No_publications") |> 
    ggplot(aes(publication_year,No_publications, color=field_1)) +
    geom_line()+
    theme_bw()  +
    theme(legend.position = "bottom") + theme(
      legend.text = element_text(size = 6),
      legend.title = element_blank(),
      legend.spacing = unit(0.05,"lines"),
      legend.key.size = unit(0.5, "lines"),
      panel.grid = element_blank())  +
    scale_x_continuous(expand = c(0,0), limits = c(1980,2025)) +
    labs(x = "Year", y="Publications") 


N_citations_over_time_field_tbl |> 
    pivot_longer(!publication_year, names_to = "field_1",values_to = "No_citations") |> 
    ggplot(aes(publication_year,No_citations, color=field_1)) +
    geom_line()+
    theme_bw()  +
    theme(legend.position = "bottom") + theme(
      legend.text = element_text(size = 6),
      legend.key.size = unit(0.5, "lines"),
      legend.title = element_blank(),
      legend.spacing = unit(0.01,"lines"),
      panel.grid = element_blank())  +
    scale_x_continuous(expand = c(0,0), limits = c(1980,2025)) +
    labs(x = "Year", y="Citations") 

```

## 4.4 Distribution of Fields across Queries


```{r query_dist, out.width='100%', fig.height=16, fig.width=16}
data_raw |> 
    filter(field_1 %in% fields_of_interest) |> 
    group_by(query, field_1) |> 
    summarise(N = n()) |>
    group_by(query) |> 
    slice_max(order_by = N, n=3, with_ties = F) |> 
    rename(Field = field_1) |> 
    ggplot(aes(x = query, y = N, fill = Field)) +
    geom_col(stat = "identity", position = "dodge") +
    labs(
      title = "Top 3 Fields per Query",
      x = "Topic",
      y = "Count"    
    ) +
    theme_minimal(base_size = 14) +
    coord_flip() 
     
```

## 4.5 Distribution of Journals across fields

```{r journal_field_dist_tbl}

journal_field_dist_raw <- data_raw |> 
    drop_na(source_display_name, field_1) |> 
    group_by(field_1, source_display_name) |> 
    summarise(N = n()) |> 
    group_by(field_1) |> 
    slice_max(order_by = N, n=3, with_ties = F) |> 
    mutate(N_total = sum(N),
           N_rel= N/N_total) |> 
    filter(N_total > 10)
  
journal_field_dist_clean <- journal_field_dist_raw |> 
    mutate(source_display_name = paste0(source_display_name, " (", N, ")")) |> 
    mutate(journal =  paste0("Journal ",row_number())) |> 
    pivot_wider(id_cols = field_1, names_from = journal, values_from = source_display_name)
  
colnames(journal_field_dist_clean)[1] <- "Field"

kable(journal_field_dist_clean, "latex", booktabs = TRUE, linesep = "", caption = "Top Journals per Field") |> 
    add_header_above(c(" " = 1, "Top Journals per Field (Descending Order)" = 3)) |> 
    column_spec(c(1:4), width = "4cm") |> 
    kable_styling(font_size = 10)


```


```{r journal_field_dist_fig, fig.height=14, fig.width=14}


journal_field_dist_raw |> 
    ggplot(aes(x = field_1, y = N_rel, fill = source_display_name)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(
      title = "Top 3 Journals per Field",
      x = "Field",
      y = "Relative Frequency"
    ) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "bottom") +
    scale_fill_discrete("Journal") +
    coord_flip()

```


# 5. Authors Overview

## 5.1 Top Authors

```{r authors, fig.cap="Top Authors in Terms of Publications or Citations"}
top_authors_publications_tbl <- data$top_authors_publications_tbl[1:30,]

top_authors_citations_tbl <- data$top_authors_citations_tbl[1:30,]

top_authors_publications_tbl |> 
    ggplot(aes(No_publications,fct_reorder(author, No_publications))) +
    geom_bar(stat = "identity", color = "black", fill = "grey") +
    labs(y="Author", x="Publications") +
    theme_bw() +
    theme(panel.grid = element_blank()) +
 scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))  # 0% below, 5% above
  )    
  
top_authors_citations_tbl |> 
    ggplot(aes(No_citations,fct_reorder(author, No_citations))) +
    geom_bar(stat = "identity", color = "black", fill = "grey") +
    labs(y="Author", x="Citations") +
    theme_bw() +
    theme(panel.grid = element_blank()) +
 scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))  # 0% below, 5% above
  )
```

## 5.2 Top Authors per Domain

```{r authors per domain}
kable(data$top_authors_domain_citations_tbl, booktabs=T, "latex", linesep="", caption = "Top Authors in Terms of Citations per Domain") |> 
  kable_styling(position = "center", latex_options = c("HOLD_position", "scale_down"), font_size = 10)

kable(data$top_authors_domain_publications_tbl, booktabs=T, "latex", linesep="", caption = "Top Authors in Terms of Publications per Domain") |> 
  kable_styling(position = "center", latex_options = c("HOLD_position", "scale_down"), font_size = 10)
```

\newpage

## 5.3 Top Authors per Field

```{r authors per field}

test <- data$top_authors_field_citations_tbl[1:3,] |> t() |> data.frame() |> rownames_to_column("Field") |> 
  filter(Field %in% fields_of_interest)
colnames(test)[2:4] <- ""

kable(test, booktabs=T, "latex", linesep="", caption = "Top Authors in Terms of Citations per Field") |> 
  add_header_above(c(" " = 1, "Authors" = 3)) |> 
  kable_styling(font_size = 8, latex_options = "HOLD_position") 

test2 <- data$top_authors_field_publications_tbl[1:3,] |> t() |> data.frame() |> rownames_to_column("Field") |> 
    filter(Field %in% fields_of_interest)

colnames(test2)[2:4] <- ""

kable(test2, booktabs=T, "latex", linesep="", caption = "Top Authors in Terms of Publications per Field") |> 
  add_header_above(c(" " = 1, "Authors" = 3)) |> 
  kable_styling(font_size = 8, latex_options = "HOLD_position")
```

\newpage

# 6. Journal Overview

```{r journals, fig.cap="Top Journals in Terms of Publications or Citations"}

top_journals_publications_tbl <- data$top_journals_publications_tbl
top_journals_citations_tbl <- data$top_journals_citations_tbl


top_journals_publications_tbl |> 
    arrange(desc(No_publications)) |> 
    slice_max(order_by = No_publications, n = 30) |> 
    ggplot(aes(No_publications,fct_reorder(journal, No_publications))) +
    geom_bar(stat = "identity", color="black", fill="grey") +
    labs(y="Journal", x="Publications") +
    theme_bw() +
    theme(panel.grid = element_blank()) +
 scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))  # 0% below, 5% above
  )  
top_journals_citations_tbl |> 
    arrange(desc(No_citations)) |> 
    slice_max(order_by = No_citations, n = 30) |> 
    ggplot(aes(No_citations,fct_reorder(journal, No_citations))) +
    geom_bar(stat = "identity", color="black", fill="grey") +
    labs(y="Journal", x="Citations") +
    theme_bw() +
    theme(panel.grid = element_blank()) +
 scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))  # 0% below, 5% above
  )
```

# 7. Distribution of Concepts

```{r concept_dist, fig.width=16, fig.height=12, out.width='100%'}

data_concepts |> 
    drop_na(level) |> 
    group_by(display_name) |> 
    summarise(N = n(), level = unique(level)) |> 
    group_by(level) |> 
    slice_max(order_by = N, n = 50) |> 
    mutate(display_name = fct_reorder(display_name, N)) |> 
    ggplot(aes(x = display_name, y = N)) +
    geom_col() +
    facet_wrap(~level, scales = "free") +
    coord_flip()

```


# 8. Keywords
```{r keywords}

keywords_tbl <- data$keywords_tbl
colnames(keywords_tbl) <- c("Word", "N")

keywords_tbl |> 
  ggplot(aes(N,fct_reorder(Word, N))) +
  geom_bar(stat = "identity", color="black", fill ="grey") +
  labs(y="Word", x="N") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
 scale_x_continuous(
    expand = expansion(mult = c(0, 0.05))  # 0% below, 5% above
  )

kable(data$keywords_tbl, booktabs=T, "latex", linesep="", caption = "Top Keywords") |> 
  kable_styling(position = "center", latex_options = "HOLD_position")

# keywords_prep <- data$keywords_prep[1:200,]
# 
# set.seed(123)  # for reproducibility
# ggplot(keywords_prep, aes(label = keywords, size=N))+
#   geom_text_wordcloud()

```

