---
title: "Field-Level Analysis"
output: pdf_document
date: "2025-11-20"
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F,
                      fig.width = 10, fig.height = 8, dpi = 300, out.width = "100%")

library(tidyverse)
library(openalexR)
library(tidyverse)
library(here)
library(bibliometrix)
library(bibliometrixData)
library(ggwordcloud)
library(kableExtra)
library(ggalluvial)
library(knitr)
library(kableExtra)
library(tidytext)


unfill_vec <- function(x) {
  same <- x == dplyr::lag(x)
  ifelse(!is.na(same) & same, "", x)
}

```


```{r data load and prep, cache=TRUE}


rds_files <- paste0("Evolution Fields/",list.files("Evolution Fields"))

rds_files_field_field <- paste0("Field_Field_Data/", list.files("Field_Field_Data"))


fields_of_interest <- c("Computer Science", "Business, Management and Accounting", "Decision Sciences", "Social Sciences",
                        "Engineering", "Economics, Econometrics and Finance")


# Read and combine
all_metrics <- rds_files %>%
  map_dfr(function(f) {
    
    
    df <- readRDS(f)
    
    # Extract window from filename
    # Assumes filenames like "1970-1974_metrics.RDS"
    window <- str_extract(basename(f), "\\d{4}-\\d{4}")
    
    df %>% mutate(window = window)
  }) |> 
  drop_na() |> 
  mutate(window = factor(window, levels = unique(window))) |> 
  group_by(window) |> 
  mutate(totalnumbernodes = sum(numbernodes),
         totalnumberedges = sum(totaledges)) |> 
  ungroup() |> 
  mutate(relative_size_nodes = numbernodes/totalnumbernodes,
         relative_size_edges = totaledges/totalnumberedges,
         relative_size_within_edges = withinedges/totalnumberedges,
         relative_size_between_edges = between_edges/totalnumberedges,
         size_share_window = numbernodes/totalnumbernodes) |> 
  filter(id %in% fields_of_interest)


all_metrics_field_field <- rds_files_field_field %>%
  map_dfr(function(f) {
    df <- readRDS(f)
    
    # Extract window from filename
    # Assumes filenames like "1970-1974_metrics.RDS"
    window <- str_extract(basename(f), "\\d{4}-\\d{4}")
    
    df %>% mutate(window = window)  |> 
      filter(field_min != field_max) |> 
      mutate(pair = paste(field_min, field_max, sep = " – "))
  }) |> 
  group_by(window) |> 
  mutate(N_edges_rel = N_edges/sum(N_edges)) |> 
  separate(window, into = c("start", "end"), remove=F, sep="-") |> 
  mutate_at(vars(start, end), as.numeric) |> 
  filter(start >= 1980) |> 
  filter(field_min %in% fields_of_interest & field_max %in% fields_of_interest)



df_long <- all_metrics %>%
  tidyr::pivot_longer(cols = c(internalratio, externalratio,degree,
                               betweeness, closeness, between_edges, withinedges,
                               numbernodes, totaledges, relative_size_nodes,
                               relative_size_edges,relative_size_within_edges,
                               relative_size_between_edges, citations),
                      names_to = "type", values_to = "value") |> 
  drop_na(field)


temporal_network_with_names <- readRDS("Temporal Networks/temporal_networks_with_names.RDS")




```

```{r prep citation flows, cache=TRUE}

data <- readRDS("data_final_1980.RDS") 

data_references <- readRDS("data_references.RDS")

## Extracting authors

authors <- data |> 
  select(oa_id, authorships, cited_by_count) |> 
  unnest(cols = c(authorships)) |> 
  select(oa_id, display_name, cited_by_count) 

## Authors as strings

authors_string <- authors |> 
  group_by(oa_id) |> 
  summarise(authors = paste(display_name, collapse = ";"), .groups = "drop") 

meta <- data |> 
  select(oa_id, title, publication_year, year_bucket_10, cited_by_count, topic_1, subfield_1, field_1, domain_1) |> 
  left_join(authors_string)

fields <- data |> 
  select(oa_id, topics) |> 
  unnest(cols = c(topics)) |> 
  mutate(type_no = paste0(type, "_", i)) |> 
  pivot_wider(id_cols = oa_id, names_from = type_no, values_from = display_name) |> 
  select(oa_id, field_1)

data_ref <- data |> 
  select(oa_id, referenced_works) |> 
  left_join(meta |> select(oa_id, publication_year, field_1)) |> 
  unnest_longer(referenced_works) |> 
  rename(oa_id_to = oa_id, oa_id_from = referenced_works, field_to = field_1) |> 
  left_join(data_references |> select(oa_id, field_1), by=c("oa_id_from"="oa_id")) |> 
  rename(field_from = field_1) 


windows_moving <- tibble(
  start = 1980:2020,
  end   = start + 4,
  window_moving = paste0(start, "-", end)
)

windows <- data |> 
  distinct(publication_year)

windows$window <- cut(
  windows$publication_year,
  breaks = c(seq(1980, 2021, by = 4), 2025),  # last bin ends just after 2025
  right  = FALSE,
  labels = c(
    paste(seq(1980, 2016, 4), seq(1984, 2020, 4), sep = "-"),
    "2021-2024"
  )
)



cit_link <- data_ref |> 
  drop_na(field_to, field_from) |> 
  group_by(publication_year, field_to, field_from) |> 
  summarise(N =n()) 

cit_link_moving <- cit_link %>%
  inner_join(
    windows_moving,
    by = character(),   # cross join
  ) %>%
  filter(publication_year >= start,
         publication_year <= end)

cit_link_moving <- cit_link_moving %>%
  group_by(window_moving, field_to, field_from) %>%
  summarise(N = sum(N), .groups = "drop") %>%
  group_by(window_moving, field_to) %>%
  mutate(N_total = sum(N),
         N_rel = N / N_total) %>%
  ungroup() |> 
  filter(field_to %in% fields_of_interest & field_from %in% fields_of_interest)

coc <- cit_link |> 
  filter(field_to != field_from) |> 
  group_by(publication_year, field_to) |> 
  summarise(N_outside = sum(N)) |> 
  rename(field = field_to)

coc_moving <- cit_link_moving |> 
  filter(field_to != field_from) |> 
  group_by(window_moving, field_to) |> 
  summarise(N_outside = sum(N)) |> 
  rename(field = field_to)

cic <- cit_link |> 
  filter(field_to == field_from) |> 
  group_by(publication_year, field_to) |> 
  summarise(N_inside = sum(N)) |> 
  rename(field = field_to)

cic_moving <- cit_link_moving |> 
  filter(field_to == field_from) |> 
  group_by(window_moving, field_to) |> 
  summarise(N_inside = sum(N)) |> 
  rename(field = field_to)

outside_give <- cit_link |> 
  filter(field_to != field_from) |> 
  group_by(publication_year, field_from) |> 
  summarise(N_give_out= sum(N)) |> 
  rename(field = field_from)


inside_give <- cit_link |> 
  filter(field_to == field_from) |> 
  group_by(publication_year, field_from) |> 
  summarise(N_give_in= sum(N)) |> 
  rename(field = field_from)


field_overview <- coc |> 
  left_join(cic) |> 
  left_join(outside_give) |>
  left_join(inside_give) |> 
  mutate_all(~replace_na(.x, 0)) |> 
  mutate(coc = N_outside/(N_outside+N_inside),
         cic = N_inside/(N_outside+N_inside),
         outside_give = N_give_out/(N_give_out+N_give_in),
         inside_give = N_give_in/(N_give_out+N_give_in)) |> 
  mutate_all(~replace_na(.x, 0))

field_overview_long <- field_overview |> 
  pivot_longer(cols = c(coc, cic, outside_give, inside_give), names_to = "variable", values_to = "fraction") |> 
  filter(field %in% fields_of_interest)


field_overview_moving <- coc_moving |> 
  left_join(cic_moving) |> 
  mutate_all(~replace_na(.x, 0)) |> 
  mutate(coc = N_outside/(N_outside+N_inside),
         cic = N_inside/(N_outside+N_inside)) |> 
  mutate_all(~replace_na(.x, 0))

field_overview_moving_long <- field_overview_moving |> 
  pivot_longer(cols = c(coc, cic), names_to = "variable", values_to = "fraction") |> 
  filter(field %in% fields_of_interest)

field_overview_long_window <- field_overview |> 
  left_join(windows) |> 
  group_by(field, window) |> 
  summarise_at(vars(coc, cic, outside_give, inside_give), mean) |> 
  pivot_longer(cols = c(coc, cic, outside_give, inside_give), names_to = "variable", values_to = "fraction") |> 
  filter(field %in% fields_of_interest) |> 
  mutate(
    window_start = as.numeric(sub("-.*$", "", window)),
    window_end   = as.numeric(sub("^.*-", "", window)),
    window_mid   = (window_start + window_end) / 2
  )


cit_link_filt <- cit_link |> 
  filter(field_to %in% fields_of_interest &  field_from %in%  fields_of_interest) |>
  group_by(field_to, publication_year) |> 
  mutate(N_total = sum(N), N_rel =N/N_total) |> 
  left_join(windows)  %>%
  mutate(
    window_start = as.numeric(sub("-.*$", "", window)),
    window_end   = as.numeric(sub("^.*-", "", window)),
    window_mid   = (window_start + window_end) / 2
  )

cit_link_filt_moving <- cit_link_moving |> 
  filter(field_to %in% fields_of_interest & field_from %in%  fields_of_interest) |>
  group_by(field_to, window_moving) |>  
  mutate(N_total = sum(N), N_rel =N/N_total)  %>%
  mutate(
    window_start = as.numeric(sub("-.*$", "", window_moving)),
    window_end   = as.numeric(sub("^.*-", "", window_moving)),
    window_mid   = (window_start + window_end) / 2
  )



```



```{r plotting funcs, cache=TRUE}
plot_type <- function(data){
  
  data_plot <- data |> 
    filter(!is.na(type)) |> 
    group_by(window, type) |> 
    summarise(value = mean(value), .groups = "drop")
  
  ggplot(data_plot, aes(x = window, y = value)) +
    geom_jitter(width = 0.2, height = 0, alpha = 0.7) +
    theme_minimal() +
    facet_wrap(~type, scales="free") +
    theme(axis.text = element_blank()) +
    labs(y = "Metric", x = "Window")
  
  
}


plot_type_field <- function(data,metric){
  
  data_plot <- data |> 
    filter(type %in% metric) |> 
    mutate(field = factor(field, levels = fields_of_interest)) |> 
    filter(!is.na(field))
  
  if(length(metric) >= 2){
    
    name <- paste(metric, collapse = "\\/")
    
    ggplot(data_plot, aes(x = window, y = value,color=type)) +
    geom_jitter(width = 0.2, height = 0, alpha = 0.7) +
    theme_minimal() +
facet_wrap(
    ~field, 
    ncol = 3,
    labeller = labeller(field = function(x) str_wrap(x, width = 30))
  ) +    theme(strip.text.x = element_text(size = 9, margin = margin(t = 5, b = 5)),
          strip.text.y = element_text(size = 9, margin = margin(l = 5, r = 5)),
          plot.margin = grid::unit(c(1,1,1,1), "cm"),
          axis.text = element_blank(),
          legend.position = "bottom") +
    labs(y = "Metric", x = "Window") 

  } else{
    
  ggplot(data_plot, aes(x = window, y = value,color=type)) +
    geom_jitter(width = 0.2, height = 0, alpha = 0.7) +
    theme_minimal() +
facet_wrap(
    ~field, 
    ncol = 3,
    labeller = labeller(field = function(x) str_wrap(x, width = 30))
  ) +    theme(strip.text.x = element_text(size = 9, margin = margin(t = 5, b = 5)),
          strip.text.y = element_text(size = 9, margin = margin(l = 5, r = 5)),
          plot.margin = grid::unit(c(1,1,1,1), "cm"),
          axis.text = element_blank(),
          legend.position = "none") +
    labs(y = toupper(metric), x = "Window")

  
  }
}


```


# 1. Aggregate Metrics
The Metrics used here are formally defined on a per-field basis below. The aggregate metrics are computed similarly, but over all fields instead of per field.

## 1.1 Evolution Aggregate of Metrics

```{r metrics_agg, cache=TRUE}
plot_type(df_long)
```

\newpage

## 1.2 Number of Nodes and Edges over Time

```{r n edges nodes}

summ <- all_metrics |> 
  group_by(window) |> 
  summarise(totalnumbernodes = unique(totalnumbernodes), totalnumberedges = unique(totalnumberedges))


kable(summ, "latex", booktabs = TRUE, linesep = "", col.names = c("Window", "Number of Nodes", "Number of Edges")) |> 
  kable_styling(position = "center", latex_options = "hold_position")


```


\newpage

# 2. Evolution of Fields

Field-Level Metrics are derived from networks where articles are nodes and edges are based on the strength of bibliographic coupling between articles. The nodes, edges and their information is then aggregated per discipline, here called field.

## 2. Relative Number of Nodes and Number of Citations

- Relative number of Nodes:
\begin{equation}
RelNodes_{t,f} = \frac{N(V)_{t,f}}{\sum_f^F N(V)_{t,f}} 
\end{equation}
where:
- $RelNodes_{t,f}$ measures the relative number of nodes for field $f$ in time window $t$ across all article nodes
- $N(V)_{t,f}$ measures the total number of nodes for field $f$ in time window $t$ across all article nodes. It is therefore defined as $N(V)_{t,f} = \sum_{v_i} 1_{field(v_i)=f}$, where $1_{field(v_i)=f}$ is an indicator function, taking the value one if article node $v_i$ belongs to field $f$.

- The number of citations per field and time window is measured as the sum of citations across all articles of a field in a give time window, respectively.

```{r nodes_citations}
plot_type_field(df_long, metric = c("relative_size_nodes")) +
  ggtitle("Evolution of Relative Number of Nodes per Field")
plot_type_field(df_long, metric = c("citations")) +
  ggtitle("Evolution of Citations per Field")




ggplot(all_metrics,
       aes(x = window,
           stratum = id,
           alluvium = id,
           y = size_share_window,
           fill = id)) +
  geom_flow(stat = "alluvium", lode.guidance = "forward", alpha = 0.8) +
  geom_stratum(width = 1/12, color = "black") +
  scale_fill_viridis_d(option = "plasma") +
  theme_minimal() +
  theme(legend.title = element_blank(), legend.position = "bottom") +
 #ggtitle("Alluvial of Field Shares over Time") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Window", y = "Field Share per Window")
```

\newpage

## 2.2 Relative Number of Edges

- Relative Number of Total Edges:
\begin{equation}
RelEdges_{t,f} = \frac{N(E)_{t,f}}{\sum_f^F N(E)_{t,f}} 
\end{equation}
where:
- $RelEdges_{t,f}$ measures the relative number of edges incident on nodes within field $f$ in time window $t$
- $N(E)_{t,f}$ measures the total number of edges incident on nodes within field $f$ in time window $t$. It is therefore defined as $N(E)_{t,f} = \sum_{e=(v_i,v_j)} 1_{field(v_i)=f \: \vee \: field(v_j) = f}$, where $1_{field(v_k)=f}$ is an indicator function, taking the value one if article node $k$ belongs to field $f$

- Relative Number of Within-Field Edges:
\begin{equation}
RelEdgesWithin_{t,f} = \frac{N(E_{within})_{t,f}}{\sum_f^F N(E)_{t,f}} 
\end{equation}
where:
- $RelEdgesWithin_{t,f}$ measures the relative number of edges between article nodes of the same field for field $f$ in time window $t$
- $N(E_{within})_{t,f}$ measures the total number of edges between article nodes of the same field for field $f$ in time window $t$. It is therefore defined as $N(E_{within})_{t,f} = \sum_{e=(v_i,v_j)} 1_{field\{(v_i)=f\}} 1_{\{field(v_j) = f\}}$, where $1_{field(v_k)=f}$ is an indicator function, taking the value one if article node $k$ belongs to field $f$.

- Relative Number of Between-Field Edges:
\begin{equation}
RelEdgesBetween_{t,f} = \frac{N(E_{Between})_{t,f}}{\sum_f^F N(E)_{t,f}} 
\end{equation}
where:
- $RelEdgesBetween_{t,f}$ measures the relative number of edges between article nodes of different fields for field $f$ in time window $t$
- $N(E_{Between})_{t,f}$ measures the total number of edges between article nodes of different fields for field $f$ in time window $t$. It's defined as $N(E_{Between})_{t,f} = \sum_{e=(v_i,v_j)} \left ( 1_{field\{(v_i=f)\}} 1_{field\{(v_j \neq f)\}} + 1_{field\{(v_i \neq f)\}} 1_{field\{(v_j = f)\}}  \right )$.


```{r edges}
plot_type_field(df_long, metric = c("relative_size_edges"))  +
  ggtitle("Evolution of Relative Number of Edges per Field") # Edges(Field)/Edges(TotalWindow)
plot_type_field(df_long, metric = c("relative_size_within_edges", "relative_size_between_edges"))  +
  ggtitle("Evolution of Relative Number of Edges within/between Fields")
# Edges(WithinField)/Edges(TotalWindow), Edges(BetweenField)/Edges(TotalWindow)
```

\newpage

## 2.3 Internal and External Edge Ratio
- Internal Edge Ratio:
\begin{equation}
IR_{t,f} = \frac{N(E_{within})_{t,f}}{N(E)_{t,f}}
\end{equation}
- External Edge Ratio:
\begin{equation}
ER_{t,f} = \frac{N(E_{between})_{t,f}}{N(E)_{t,f}}
\end{equation}

with $N(E_{within})_{t,f}$, $N(E_{between})_{t,f}$ and N(E)_{t,f} defined as before. The internal and external ratios therefore measure the ratio of edges within (between) fields relative to the total number of edges of a field. Edges are derived from article-level nodes.

```{r ratios}
plot_type_field(df_long, metric = c("internalratio", "externalratio")) +
  ylab(label = "Internal/External Edge Ratio") + scale_color_discrete(
  name = "Metric",   # legend title
  labels = c("internalratio" = "Internal Ratio", 
             "externalratio" = "External Ratio")
)   +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(breaks = c("1980-1984", "2000-2004", "2020-2024"))

# Edges(WithinField)/Edges(WithinField)+Edges(BetweenField), Edges(BetweenField)/Edges(WithinField)+Edges(BetweenField)
```

\newpage

## 2.4 Network Centrality Metrics (Degree, Closeness and Betweennes)

```{r network_stats}
plot_type_field(df_long, metric = c("degree"))  +
  ggtitle("Evolution of Field-level Degree")
```

\newpage

```{r network_stats_cont}
plot_type_field(df_long, metric = c("closeness"))  +
  ggtitle("Evolution of Field-level Closeness")
plot_type_field(df_long, metric = c("betweeness"))  +
  ggtitle("Evolution of Field-level Betweeness")
```


# 3. Field-Field Connectivity

## 3.1 Average Connectivity

```{r average_connectivity}

all_metrics_field_field_avg <- all_metrics_field_field |> 
  group_by(field_min, field_max) |> 
  summarise_at(vars(N_edges, N_edges_rel), mean)




ggplot(all_metrics_field_field_avg, aes(x = field_min, y = field_max, fill = N_edges)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "plasma") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    panel.grid = element_blank()
  ) +
  ggtitle("Average Number of Egdes between Fields")


```

\newpage

```{r average_connectivity_cont}



ggplot(all_metrics_field_field_avg, aes(x = field_min, y = field_max, fill = N_edges_rel)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "plasma") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    panel.grid = element_blank()
  ) +
  ggtitle("Average Number of Egdes between Fields (on a relative basis)")



```
\newpage

## 3.2 Connectivity over time 

### 3.2.1 Heatmaps

```{r heatmap n over time, fig.height=16, fig.width=12}

ggplot(all_metrics_field_field, aes(x = field_min, y = field_max, fill = N_edges)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "plasma") +
  facet_wrap(~ window, ncol=6) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    panel.grid = element_blank()
  ) +
  ggtitle("Number of Egdes between Fields over Time")


```


```{r heatmap n_rel over time,  fig.height=16, fig.width=12}

ggplot(all_metrics_field_field, aes(x = field_min, y = field_max, fill = N_edges)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "plasma") +
  facet_wrap(~ window, ncol=6) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1),
    panel.grid = element_blank()
  ) +
  ggtitle("Number of Egdes between Fields over Time (on a relative basis)")
 

```

### 3.2.2 Time Series

```{r edges over time, fig.height=14}
ggplot(all_metrics_field_field, aes(x = window, y = N_edges, group = pair, color = pair)) +
  geom_line() +
  geom_point(size = 1) +
  facet_wrap(~ str_wrap(pair, width = 20)) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_blank()) +
  ggtitle("Number of Egdes between Fields over Time")

```

\newpage

```{r relative edges over time , fig.height=14}
ggplot(all_metrics_field_field, aes(x = window, y = N_edges_rel, group = pair, color = pair)) +
  geom_line() +
  geom_point(size = 1) +
  facet_wrap(~ str_wrap(pair, width = 20)) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_blank())  +
  ggtitle("Number of Egdes between Fields over Time (on a relative basis)")


```

# 4. Citation Flows

## 4.1 Internal Coverage
```{r internal coverage, results='asis'}

internal_coverage <- data_ref |> 
  group_by(field_from) |> 
  summarise(N = n()) |> 
  ungroup() |> 
  mutate(N_total = sum(N), rel_coverage = round((N/N_total)*100,2)) |> 
  select(Field = field_from, "Total Number of References" = N, "Relative Number of References (in %)" = rel_coverage)

kable(internal_coverage, "latex", booktabs = TRUE, linesep="", caption = "Number of References with Meta Data") |> 
  kable_styling(latex_options = c("hold_position")) |> 
  column_spec(1, width = "5cm")

```


## 4.2 Outside/Inside Citations

### 4.2.1 Per Year

```{r outside inside year}

field_overview_long |>
  filter(variable %in% c("cic", "coc")) |> 
  ggplot(aes(x = publication_year, y = fraction, fill = variable)) +
  geom_area(alpha = 0.8) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Fraction", x = "Year", title = "Outside/Inside Citation per Year") +
  theme_minimal() +
  facet_wrap(~field) + scale_fill_discrete(
  name = "Metric",   # legend title
  labels = c("cic" = "Citations Inside Category", 
             "coc" = "Citations Oustide Category")
) +
  theme(legend.position = "bottom") 



```

\newpage

```{r outside inside year cont}



field_overview_long |>
  filter(variable %in% c("outside_give", "inside_give")) |> 
  ggplot(aes(x = publication_year, y = fraction, fill = variable)) +
  geom_area(alpha = 0.8) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Fraction", x = "Year", title = "Outside/Inside Reference per Year") +
  theme_minimal() +
  facet_wrap(~field) + scale_fill_discrete(
  name = "Metric",   # legend title
  labels = c("inside_give" = "Reference Inside Category", 
             "outside_give" = "Reference Oustide Category")
) +
  theme(legend.position = "bottom")

```

### 4.2.2 Per Window (non-overlapping)

```{r outside inside window}
field_overview_long_window |>
  filter(variable %in% c("cic", "coc")) |> 
  ggplot(aes(x = window_mid, y = fraction, fill = variable)) +
  geom_area(alpha = 0.8) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Fraction", x = "Window", title = "Outside/Inside Citation per Window") +
  theme_minimal() +
  facet_wrap(~field) + scale_fill_discrete(
  name = "Metric",   # legend title
  labels = c("cic" = "Citations Inside Category", 
             "coc" = "Citations Oustide Category")
) +
  theme(legend.position = "bottom")


```

\newpage

```{r outside inside window cont}


field_overview_long_window |>
  filter(variable %in% c("outside_give", "inside_give")) |> 
  ggplot(aes(x = window_mid, y = fraction, fill = variable)) +
  geom_area(alpha = 0.8) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Fraction", x = "Window", title = "Outside/Inside Reference per Window") +
  theme_minimal() +
  facet_wrap(~field) + scale_fill_discrete(
  name = "Metric",   # legend title
  labels = c("inside_give" = "Reference Inside Category", 
             "outside_give" = "Reference Oustide Category")
) +
  theme(legend.position = "bottom")
```

### 4.2.3 Per Window (Overlapping)

```{r outside inside window overlapping}
field_overview_moving_long |>
  filter(variable %in% c("cic", "coc")) |> 
  ggplot(aes(x = window_moving, y = fraction, fill = variable)) +
  geom_area(aes(group = variable),alpha = 0.8) +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Fraction", x = "Window") +
  theme_minimal() +
  facet_wrap(~field) + scale_fill_discrete(
  name = "Metric",   # legend title
  labels = c("cic" = "Citations Inside Category", 
             "coc" = "Citations Oustide Category")
) +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(breaks = c("1980-1984", "2000-2004", "2020-2024"))



```


## 4.3 Citation Flows over Time


```{r citation over time}

ggplot(cit_link_filt, aes(x = publication_year, y = N_rel, color = field_from)) +
  geom_line() +
  facet_wrap(~ field_to) +
  theme_minimal() +
  scale_color_discrete(name = "Field", guide = guide_legend(nrow=2))  +
  theme(legend.position = "bottom") +
  labs(y = "Fraction of Citations from another Field", x = "Year", title = "Fraction of Citations from a Field per Year")

```

\newpage

```{r citation over time cont}
  
  
ggplot(cit_link_filt_moving, aes(x = window_mid, y = N_rel, color = field_from)) +
  geom_line() +
  facet_wrap(~ field_to) +
  theme_minimal()+
  scale_color_discrete(name = "Field", guide = guide_legend(nrow=2))  +
  theme(legend.position = "bottom") +
  labs(y = "Fraction of Citations from another Field", x = "Year", title = "Fraction of Citations from a Field per 5-year Rolling Windows")


```


## 4.4 Alluvial Graph of Flows

```{r alluvial, fig.height=12, fig.width=16}

cit_link_filt_window <- cit_link_moving |> 
  filter(window_moving %in% c("1980-1984", "2000-2004", "2020-2024")) 
  

ggplot(cit_link_filt_window,
       aes(axis1 = field_from, axis2 = field_to, y = N_rel)) +
  geom_alluvium(aes(fill = field_from), width = 1/12, alpha = 0.8) +
  geom_stratum(width = 1/4, fill = "grey80", color = "black") +
  geom_text(stat = "stratum", 
            aes(label = str_wrap(after_stat(stratum), width = 12))) +  
  facet_wrap(~ window_moving, ncol = 3, scales = "free_y") +
  labs(y = "Fraction of outgoing citations",
       x = NULL) +
  theme_minimal()+
  theme(legend.position = "none")


```

\newpage

## 4.5 Citation Flow Dispersion

```{r citation disp}

disp_overall <- cit_link_moving |> 
  group_by(field_to) |> 
  summarise(entropy = sum(-N_rel*log(N_rel)))

kable(disp_overall, "latex", booktabs = T, linesep = "", col.names = c("Field", "Flow Entropy"), caption = "Citation Flow Entropy across Fields") |> 
  kable_styling(latex_options = "hold_position")


```

\newpage

```{r citation disp cont}


disp_window <- cit_link_moving |> 
  group_by(field_to, window_moving) |> 
  summarise(entropy = sum(-N_rel*log(N_rel)))
  


disp_window |> 
  ggplot(aes(window_moving, entropy)) +
  geom_point() +
  facet_wrap(~field_to) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_discrete(breaks = c("1980-1984", "2000-2004", "2020-2024")) +
  labs(x = "Window", y = "Entropy", title = "Field-level Entropy over Time")

```




\newpage

# 5. Topical Analysis of Fields

```{r topic_overview}

data_topics <- readRDS("field_topic_data.RDS") |> 
  select(-text, -embedding) |> 
  filter(topic != -1) |> 
  mutate(topic_keywords = str_replace_all(topic_keywords, "\\[", ""),
         topic_keywords = str_replace_all(topic_keywords, "\\]", ""),
         topic_keywords = str_replace_all(topic_keywords, "\\'", ""))


topic_overview <-  data_topics |> 
  group_by(field_1, topic, topic_keywords) |> 
  summarise(N = n()) |> 
  mutate(field_1 = unfill_vec(field_1)) 

kbl_obj <- kable(topic_overview, "latex", booktabs = TRUE, linesep="", caption = "Overview of Topics within Fields.", longtable = T, 
                 col.names = c("Field", "Topic No.", "Keywords","No. Articles"), align = "lclc") |> 
  kable_styling(latex_options = c("hold_position")) |> 
  column_spec(1, width = "4cm", border_right = T) |> 
  column_spec(2, width = "2cm", border_right = T) |> 
  column_spec(3, width = "8cm", border_right = T) |> 
  column_spec(4, width = "2cm")

num_rows <- nrow(topic_overview)
for (i in seq(9, num_rows, by = 9)) {
  kbl_obj <- kbl_obj |> row_spec(i, extra_latex_after = "\\midrule")
}

kbl_obj

```

\newpage

```{r topic_overview_cont, fig.height=12}

data_topics |> 
  group_by(field_1, topic, topic_keywords) |> 
  summarise(N = n()) |> 
  group_by(field_1) |> 
  mutate(N_rel = N / sum(N)) |> 
  slice_max(order_by = N_rel, n = 3, with_ties = FALSE) |>
  ungroup() |>
  ggplot(aes(
    x = reorder_within(topic_keywords, N_rel, field_1),
    y = N_rel,
    fill = topic_keywords
  )) +
  geom_col() +
  labs(x = "") +
  scale_x_reordered() +
  facet_wrap(~ field_1, scales = "free_x")  +
  labs(x = "") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.position = "bottom", legend.direction = "vertical",
        legend.title = element_text(hjust = 0.5),
    legend.key.width = unit(0.5, "cm"),            
    legend.box = "vertical",                      
    legend.box.margin = margin(5, 5, 5, 5),
    legend.text = element_text(size = 7)) +
  scale_fill_discrete("Keywords", guide = guide_legend(ncol = 2))

```

